Adversarial Training of Neural Networks\\ against Systematic Uncertainty

### 2016-08-02 paper.tex
save

Let assume a probability space $(\Omega, {\cal F}, P)$, where $\Omega$ is a
sample space, ${\cal F}$ is a set of events and $P$ is a probability measure.
Let consider the multivariate random variables $X_\lambda: \Omega \mapsto \mathbb{R}^p$
and $Y: \Omega \mapsto {\cal Y}$, where $X_\lambda$ depends on a nuisance
parameter $\lambda$ encoding for systematic uncertainties in its probability
distribution. That is, $X_\lambda$ and $Y$ induce together the joint probability
distribution $p(X_\lambda,Y)$. For training, let further assume a finite set $\{
 \lambda_i, x_i, y_i, \}_{i=1}^N$ of realizations $X_{\lambda_i}(\omega_i),
Y(\omega_i)$, for $\omega_i \in \Omega$ and known values $\lambda_i$ of the
nuisance parameter. Our goal is to learn a function $f : \mathbb{R}^p \mapsto
{\cal Y}$ (e.g., a classifier if ${\cal Y}$ is a finite set of classes)
minimizing the expected value of a loss $L(Y, f(X_{\lambda}))$, with the constraint that
$f(X_\lambda)$ should be robust to the value of the nuisance parameter
$\lambda$ -- which remains unknown at test time. More specifically, we aim at building $f$ such that in the ideal case
\begin{equation}\label{eqn1}
f(X_{\lambda_i}(\omega)) = f(X_{\lambda_j}(\omega))
\end{equation}
for any sample $\omega \in
\Omega$ and any $\lambda_i, \lambda_j$ pair of values of the nuisance parameter.

Since we do not have training tuples $(X_{\lambda_i}(\omega),
X_{\lambda_j}(\omega))$ (for the same unknown $\omega$), we propose instead to
solve the closely related problem of finding a predictive function $f$ such that
\begin{equation}\label{eqn2}
    P(\{ \omega | f(X_{\lambda_i}(\omega)) = y \} ) = P( \{ \omega' | f(X_{\lambda_j}(\omega')) = y\}) \text{ for all $y \in {\cal Y}$}.
\end{equation}
In words, we are looking for a predictive function $f$ such that  the
distribution of $f(X_\lambda)$ is invariant with respect to the nuisance
parameter $\lambda$. Note that a function $f$ for which Eqn.~\ref{eqn1} is
true necessarily satisfies Eqn.~\ref{eqn2}. The converse is however in
general not true, since the sets of samples $\{ \omega | f(X_{\lambda_i}(\omega)) = y \}$
and $\{ \omega' | f(X_{\lambda_j}(\omega)) = y \}$ do not need to be the same
for the equality to hold. \glnote{Yet, in practice, this criterion is often good enough.}

>>>>

Let assume a probability space $(\Omega, {\cal F}, P)$, where $\Omega$ is a
sample space, ${\cal F}$ is a set of events and $P$ is a probability measure.
Let consider the multivariate random variables $X_\lambda: \Omega \mapsto
\mathbb{R}^p$ and $Y: \Omega \mapsto {\cal Y}$, where $X_\lambda$ depends on a
nuisance parameter $\lambda$ whose values define the family of its systematic uncertainties.
That is, $X_\lambda$ and $Y$ induce together
a joint probability distribution written as $p(X,Y|\lambda)$, where the
conditional on $\lambda$ denotes $X_\lambda$. For training, let further assume a
finite set $\{ \lambda_i, x_i, y_i, \}_{i=1}^N$ of realizations
$X_{\lambda_i}(\omega_i), Y(\omega_i)$, for $\omega_i \in \Omega$ and known
values $\lambda_i$ of the nuisance parameter. Our goal is to learn a function $f :
\mathbb{R}^p \mapsto {\cal Y}$ (e.g., a classifier if ${\cal Y}$ is a finite set
of classes) minimizing the expected value of a loss $L(Y, f(X_{\lambda}))$, with
the constraint that $f(X_\lambda)$ should be robust to the value of the nuisance
parameter $\lambda$ -- which remains unknown at test time. More specifically, we
aim at building $f$ such that in the ideal case \begin{equation}\label{eqn1}
f(X_{\lambda_i}(\omega)) = f(X_{\lambda_j}(\omega)) \end{equation} for any
sample $\omega \in \Omega$ and any $\lambda_i, \lambda_j$ pair of values of the
nuisance parameter.

Since we do not have training tuples $(X_{\lambda_i}(\omega),
X_{\lambda_j}(\omega))$ (for the same unknown $\omega$), we propose instead to
solve the closely related problem of finding a predictive function $f$ such that
\begin{equation}\label{eqn2}
    P(\{ \omega | f(X_{\lambda_i}(\omega)) = y \} ) = P( \{ \omega' | f(X_{\lambda_j}(\omega')) = y\}) \text{ for all $y \in {\cal Y}$}.
\end{equation}
In words, we are looking for a predictive function $f$ such that  the
distribution of $f(X_\lambda)$ is invariant with respect to the nuisance
parameter $\lambda$. Note that a function $f$ for which Eqn.~\ref{eqn1} is
true necessarily satisfies Eqn.~\ref{eqn2}. The converse is however in
general not true, since the sets of samples $\{ \omega | f(X_{\lambda_i}(\omega)) = y \}$
and $\{ \omega' | f(X_{\lambda_j}(\omega')) = y \}$ do not need to be the same
for the equality to hold. \glnote{This criterion is still adequate for most purposes.}
